---
title: "Normalization and Transformation"
output: html_document
---

To estimate a correction for sequencing depth and perform simple EDA of the samples by cell type.

## Data

```{r}
download.file("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/wang_eset.RData", "wang_eset.RData")
load("wang_eset.RData")
```

Take a subset of the count matrix and the column data for building a DESeqDataSet:

```{r message=FALSE,warning=FALSE}
library(Biobase)
count.matrix <- exprs(wang.eset)[,10:21]
col.data <- pData(wang.eset)[10:21,]
library(DESeq2)
dds <- DESeqDataSetFromMatrix(count.matrix, col.data, design=~cell.type)
dds$cell.type
```

Levels of cell.type with no associated samples are dropped.



## Q1. What tissue has the highest size factor?

```{r}
sizeFac <- estimateSizeFactors(dds)
sizeFac$cell.type[which.max(sizeFac$sizeFactor)]
```



## Q2. What sample clusters with cerebellum? 

Run the varianceStabilizingTransformation() transformation (blind=FALSE) on the counts, save this to an object 'vsd', and then make a PCA plot with "cell.type" as the color of the points.

```{r}
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
plotPCA(vsd, intgroup="cell.type")
```



## Q3. Which tissue in 'mat' has the highest correlation to skeletal muscle? 

Make a scatterplot matrix of some of the samples' transformed counts.

```{r}
rmeans <- rowMeans(assay(vsd)) # row mean of rlog-transformed data
idx <- c(1,2,10,7,8,9,12) # pick some samples for visualization
mat <- assay(vsd)[rmeans > 1,idx] # pull out a small matrix of rlog-transformed counts
colnames(mat) <- vsd$cell.type[idx] # name the columns of matrix by cell type

panel.sub <- function(x,y,...) points(cbind(x,y)[sample(length(x),1000),],...)
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)  {
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(mat, asp=1, col=rgb(0,0,0,.3), lower.panel=panel.cor, upper.panel=panel.sub)

colnames(mat[,-7])[ which.max(cor(mat[,7], mat[,-7])) ]
```

